---
title: "MNIST Example"
output: rmarkdown::github_document
---

#### Calculate the Combinatorial Laplacian score of pixels on a nerve complex created by TDA mapper run on the MNIST dataset

```{r}
library(RayleighSelection)
```

## Load the mnist dataset
```{r}
data("mnist")
```

## Compute reduced representation
#### Using Laplacian eigenmap of pixels with high variance
```{r}
library(dimRed)
leim <- LaplacianEigenmaps()
mnist_top <- mnist[apply(mnist, 1, var) > 10000,]
emb <- leim@fun(as(t(mnist_top), "dimRedData"), leim@stdpars)
```

## Compute Mapper representation
#### Using the Laplacian eigenmap as an auxiliary function and correlation distance as metric
```{r echo = T, results = 'hide'}
library(TDAmapper)
mnist_distances <- (1.0 - cor(mnist_top))
m2 <- mapper2D(distance_matrix = mnist_distances,
                filter_values = list(emb@data@data[,1], emb@data@data[,2]),
                num_intervals = c(30,30),
                percent_overlap = 35,
                num_bins_when_clustering = 10);
```

## Compute the nerve complex
```{r}
gg <- nerve_complex(m2$points_in_vertex)
```

## Compute 0-form and 1-form Comb. Lap. scores, p-value, and q-value
#### For the 301st through 305th pixels
```{r}
rayleigh_selection(gg, mnist[301:305,])
```
