---
title: "Vietoris-Rips Complex Example"
output: rmarkdown::github_document
---

#### Calculate the Combinatorial Laplacian score of genes on a VR complex from the first 50 principal components of gene expression data from two cell differentiation paths of mouse embryonic cells

```{r}
library(RayleighSelection)
```

## Load data
```{r}
expression <- read.table(file = "https://www.dropbox.com/s/3ohj3evv3zwzrs9/filtered_normalized_counts_ordered.csv?dl=1", sep=",", header=TRUE, row.names=1, stringsAsFactors=FALSE)

pca <- read.table("https://www.dropbox.com/s/onmnzlerl56ckq9/pca_50_ordered.csv?dl=1", sep=",", header=TRUE, row.names=1, stringsAsFactors = FALSE)
```

### Subsample data
#### Take a random sample of cells so L1 laplacian runs faster
```{r}
# A subsample of 100 cells is loaded below - reduce the number of cells to run faster
# subsample_cells <- sample(nrow(pca), 100, replace=FALSE)
# subsample_cells <- subsample_cells[order(subsample_cells)]

# Use these cells if you want to directly reproduce the R0 and R1 scores below
subsample_cells <- scan(file="https://www.dropbox.com/s/84z0poyz2vhqis8/tutorial_subsample.txt?dl=1", what=numeric())
```

```{r}
subsample_pca <- pca[subsample_cells,]
subsample_expression <- expression[,subsample_cells]

# Take only genes that occur in >5% and <50% of cells
subsample_expression <- subsample_expression[rowSums(subsample_expression != 0)>5,]
subsample_expression <- subsample_expression[rowSums(subsample_expression != 0)<50,]
```

## Create Euclidean distance matrix from principal components
```{r}
distance_matrix <- as.matrix(dist(subsample_pca, method="euclidean"))
```



# Only 0-form Combinatorial Laplacian score

## Create the 1-skeleton of a Vietoris-Rips complex from the distance matrix
#### With radius=28
```{r}
gg <- vr_complex(distance_matrix, 28, clique = FALSE)
```

### Show Vietoris-Rips complex
```{r}
plot_skeleton(gg)
```

## Compute 0-form Comb. Lap. score, p-value, and q-value 
#### For all genes in the expression matrix
#### Only use 1-skeleton for 0th Comb. Lap. score
```{r}
scoresR0 <- rayleigh_selection(gg, subsample_expression, num_perms = 1000, num_cores = 8, one_forms = FALSE)
(scoresR0[order(scoresR0$R0),])[1:5,]
```



# Both 0-form and 1-form Combinatorial Laplacian score

## Create the Vietoris-Rips complex from the distance matrix on first 20 cells
#### Including higher-order relations
#### With radius=28
```{r}
gg <- vr_complex(distance_matrix, 28)
```

### Show Vietoris-Rips complex
```{r}
plot_skeleton(gg)
```

## Compute 0-form and 1-form Comb. Lap. scores, p-value, and q-value 
#### For all genes in the expression matrix
This will take around 1 hour to run. Decrease the number of cells to run faster.
```{r}
scores <- rayleigh_selection(gg, subsample_expression, num_perms = 1000, num_cores = 8)
(scores[order(scores$R1, scores$R0),])[1:5,]
```
